---
title: "Report"
date: "11/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Start of code (place holder until we have page)
```{r loading packages, echo = F, message = F}
library(tidyverse)
library(viridis)
library(readxl)
library(plotly)
library(modelr)

df = 
  read_csv("./data/squirrel_data.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  plot_ly(
  x = ~x, y = ~y, type = "scatter", mode = "markers")
```
## Group members (names and unis)
Alexis Colangelo, avc2129;
Dania Jafar, dj2536;
Si Li, sl4657;
Kristal Quispe, kq2127;
Barik Rajpal, bsr2136

## Regression Analysis

It is likely that squirrels exhibit different behaviors in the morning (AM) and evening (PM) due to sleep, eating, and predator habits. We can examine if squirrel behaviors can allow us to predict whether a sighting occured during the AM or PM using a logistic regression model. We will tidy the dataset to include only squirrel behaviors, the am/pm response variable, and an ID to partition the data. We'll then split the data into a training and testing set to test multiple models.

```{r lm analysis start, message = F}
sq <- read_csv("./data/squirrel_data.csv") %>%
 mutate(AM = ifelse(Shift=="AM",1,0),
        id = row_number()) %>%
  janitor::clean_names() %>%
  select(id, am, running, chasing, climbing, eating, foraging, kuks, quaas, moans, tail_flags, tail_twitches,approaches,indifferent,runs_from)

set.seed(100)
sq_train <- sample_frac(sq, .8)
sq_test <- anti_join(sq,sq_train, by = "id")
```

We will create a glm using the step function, with the direction = "backward". This will start by creating a regression model with all the potential covariates, and drop the least significant one until it is only left with significant covariates. We are not showing the process of the step function (as it can be quite long), but we will show a summary of the model created.

```{r creating a model, echo = F, results = F}
mod1 <- step(glm(am ~., data = sq_train, family = "binomial"),direction = "backward")
```

```{r summary of mod 1}
summary(mod1)
```
The step function left many covariates in, but since they are all binary and some behaviors are not displayed very often, we believe it is likely that the model has too many covariates, and could be reduced further. We will also try one model using only the covariates with high degrees of significance, climbing, eating, approaches, and indifferent. We will then compare the two models.

``` {r creating a second model}  
mod2 <- glm(formula = am ~ climbing + eating + approaches +
    indifferent, family = "binomial", data = sq_train)

summary(mod2)

anova(mod2,mod1,test="Chisq")
```

In both models, eating as a squirrel behavior is more associated with the PM, and climbing, indifferent, and approaches behaviors are more associated with the AM.

The AIC increased slightly when removing three covariates, but only slightly, so the smaller model may be better due to parsimony. The ANOVA however, did tell us that the model with more covariates was statistically significant. We can further evaluate the results of the models using a 2x2 table of accuracy in predictions in the test dataset.

```{r evaluating the models, echo = F}
sq_test %>%
  mutate(pred = predict.glm(mod1, sq_test),
         pred_cleaned = ifelse(pred>=0,1,0)) %>%
  group_by(am,pred_cleaned) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = pred_cleaned,
              values_from = n) %>%
  knitr::kable(caption = "Accuracy for mod1")

sq_test %>%
  mutate(pred = predict.glm(mod2, sq_test),
         pred_cleaned = ifelse(pred>=0,1,0)) %>%
  group_by(am,pred_cleaned) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = pred_cleaned,
              values_from = n) %>%
  knitr::kable(caption = "Accuracy for mod2")
```

Model 1 with more covariates had better specificity 281/(281+52) = 84%, but very poor sensitivity 67/(67+205) = 24.5%. Model 2 with fewer covariate had a slightly lower specificity 237/(237+96) = 71%, and slightly better sensitivity 118/(118+154) = 43%.

Regarding overall accuracy, model 2 was slightly better, but both models predictions are quite poor (below 60% accuracy for both), so it does not appear to be reasonable to attempt to predict the time of day based on observed squirrel behavior. Part of this may be due to the limited amount of squirrel behaviors recorded per row.

```{r squirrel beaviors per row, echo = F}
sq %>% 
  mutate(total_behaviors = running + chasing + climbing + eating + foraging + kuks + quaas + moans + tail_flags + tail_twitches + approaches + indifferent + runs_from) %>%
  select(total_behaviors) %>%
  summary()
```

With a median of only 2 behaviors recorded per row, it is possible we would have been able to create a better logisitic regression model if the squirrels were observed for a longer period of time per recorded observation, so that many behaviors could be observed and recorded at once.

## End of Regression Analysis

The written report produced by your team is central to this project. This will detail how you completed your project, and should cover data collection and cleaning, exploratory analyses, alternative strategies, descriptions of approaches, and a discussion of results. We anticipate that your project will change somewhat over time; these changes and the reasons for them should be documented! You will write one report document per group, and be sure to include all group member names in the document.

Your report should include the following topics. Depending on your project type the amount of discussion you devote to each of them will vary:

* Motivation: Provide an overview of the project goals and motivation.

* Related work: Anything that inspired you, such as a paper, a web site, or something we discussed in class.

* Initial questions: What questions are you trying to answer? How did these questions evolve over the course of the project? What new questions did you consider in the course of your analysis?:

-- What is the probability of a certain squirrel in a certain area; visualize this probability/heat map
-- regression question: number (count) of squirrels seen eating as predicted by foraging 
-- Where are squirrels are most concentrated?
-- What are the most common squirrel behaviors? 

* Data: Source, scraping method, cleaning, etc.

* Exploratory analysis: Visualizations, summaries, and exploratory statistical analyses. Justify the steps you took, and show any major changes to your ideas.

* Additional analysis: If you undertake formal statistical analyses, describe these in detail
Discussion: What were your findings? Are they what you expect? What insights into the data can you make?
As this will be your only chance to describe your project in detail, make sure that your report is a standalone document that fully describes your process and results. We also expect you to write high-quality code that is understandable to an outside reader. Coding collaboratively and actively reviewing code within the team will help with this